---
title: "mylab3"
author: "Nathaniel Gunter"
date: "1/27/2020"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: spacelab
    highlights: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(s20x)
library(ggplot2)
```

## Task 1

```{r workingDir}
getwd()
```

## Task 2
```{r readDat}
spruce.df <- read.csv("SPRUCE.csv")
head(spruce.df)
```

## Task 3

```{r spruceScatter}
with(spruce.df,
  plot(Height~BHDiameter,bg="Blue",pch=21,cex=1.2, main = "Height vs BHDiameter", ylim=c(0,1.1*max(Height)),xlim=c(0,1.1*max(BHDiameter))))
```

This plot does not appear at first glance to have a straight-line relationship.

```{r trendScatter}
with(spruce.df,{layout(matrix(1:3,nr=3))
  trendscatter(x = BHDiameter,y = Height,f = 0.5)
  trendscatter(x = BHDiameter,y = Height,f=0.6)
  trendscatter(x=BHDiameter,y=Height,f=0.7)
})
```

```{r makeLinearModel}
spruce.lm<-with(spruce.df, lm(Height~BHDiameter))
with(spruce.df,
  plot(Height~BHDiameter,bg="Blue",pch=21,cex=1.2, main = "Height vs BHDiameter", ylim=c(0,1.1*max(Height)),xlim=c(0,1.1*max(BHDiameter))))
abline(spruce.lm)
```

As the lowess smoother plots show, a straight line is not appropriate for this data set. A better fit would likely be more like a $\frac{1}{x}$ relationship.

##Task 4

```{r fourbyfourspread}
layout(matrix(1:4,nr=2,nc=2,byrow=TRUE))
layout.show(4)
with(spruce.df,
     plot(Height~BHDiameter, bg="Blue",pch=21,cex=1.2,ylim=c(0,1.1*max(Height)),xlim=c(0,1.1*max(BHDiameter))))
abline(spruce.lm)

yhat=fitted(spruce.lm)

with(spruce.df, plot(Height~BHDiameter, bg = "Blue", pch=21, cex=1.2, ylim = c(0,1.1*max(Height)),xlim=c(0,1.1*max(BHDiameter))))
with(spruce.df, {
  segments(BHDiameter, Height, BHDiameter, yhat)
})
abline(spruce.lm)
with(spruce.df,
     plot(Height~BHDiameter, bg = "Blue", pch=21, cex=1.2, ylim = c(0,1.1*max(Height)),xlim=c(0,1.1*max(BHDiameter))))
with(spruce.df,abline(h=mean(Height)))     
with(spruce.df,
     segments(BHDiameter,mean(Height),BHDiameter,yhat,col="Red"))
abline(spruce.lm)
with(spruce.df,
     plot(Height~BHDiameter, bg = "Blue", pch=21, cex=1.2, ylim = c(0,1.1*max(Height)),xlim=c(0,1.1*max(BHDiameter))))
with(spruce.df, abline(h=mean(Height)))
with(spruce.df,segments(BHDiameter, Height, BHDiameter, mean(Height),col="Green"))
```

```{r ssCalc}
RSS = with(spruce.df, sum((Height-yhat)^2))
MSS = with(spruce.df, sum((yhat-mean(Height))^2))
TSS = with(spruce.df,sum((Height-mean(Height))^2))
RSS
MSS
TSS
MSS/TSS
MSS+RSS
```

The value $\frac{MSS}{TSS}$ represents TODO HERE. TSS does equal MSS+RSS.

## Task 5
```{r summarizeModel}
summary(spruce.lm)
```

The slope has the value 0.48147, and the intercept has the value 9.14684. Therefore, the equation of the line is $y=9.14684+0.48147\times x$.


```{r heightPredict}
predict(spruce.lm, data.frame(BHDiameter=c(15,18,20)))
```

## Task 6

```{r plotBuild}
g = ggplot(data=spruce.df,aes(x=BHDiameter,y=Height,colour=BHDiameter))
g = g+geom_point()+geom_line()+geom_smooth(method="lm")
g = g+labs(title="Height vs BHDiameter")
g = g+theme(plot.title = element_text(hjust = 0.5))
g
```

## Task 7

